---
title: "Homework 8"
author: "Conor Craig"
date: "2023-10-19"
output:
  pdf_document:
    latex_engine: xelatex
Header-includes:
  - \usepackage{hyperref}
  - \usepackage {float}
biblio-style: apsr
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: single
endnote: no
---

# Creating a data frame object --------------------------------------------

# Create a vector called ID that contains a number from 1 through 6

```{r}
ID <- sample (1:6)
```

# Create a vector called Race that contains the following:
# "black", "white", "black", "hispanic", "white", "white"

```{r}
race <- c("Black", "White", "Black", "hispanic", "White", "White")
```

# What is the type of the Race vector? 
# Is it a factor? Is it a character? Is it a logical vector?
# Find out by using some commands.

```{r}
# Three commands are used since it asked for "some" commands
class(race)
typeof(race)
str(race)
```

# Create a vector called Voted_For_Obama that contains thefollowing: 
# TRUE, TRUE, TRUE, FALSE, FALSE, TRUE

```{r}
voted_for_obama <- c(TRUE, TRUE, TRUE, FALSE, FALSE, TRUE) 
```

# Find out the data type of the Voted_For_Obama vector

```{r}
class(voted_for_obama)
```


# Create a vector called Party_ID that contains the following: 
# "Dem", "Dem", "Dem", "Rep", "Ind", "Rep"

```{r}
party_id <- c("dem", "dem", "dem", "rep", "ind", "rep") 
```

# Create a vector called Income_Level that contains the following: 
# "High", "Low", "Low", "High", "High", "Low"

```{r}
income_level <- c("high", "low", "low", "high", "high", "low") 
```

# Create a vector called Approval that contains the following:
# 70, 80, 68, 20, 10, 60

```{r}
approval <- c(70, 80, 68, 20, 10, 60) 
```

# Create a data frame called vote.data by combining the six vectors
# you have created above. 

```{r}
voter_data <- data.frame(ID, race, voted_for_obama, party_id, income_level, approval,
                         stringsAsFactors = TRUE)
```

# We have seen above that the Race vector was a character vector.
# We have learned in the joint exercise that we should treat a 
# vector like this one as a factor, not as a character. 

# A nice thing about data frame objects is that, it will 
# convert vectors like the Race vector into factors automatically. 
# To convince yourself that the Race vector included in the vote.data
# is indeed a factor vector, apply the is.factor function on it. 

```{r}
is.factor(voter_data$race) 
```

# Notice that is.factor(Race) returns FALSE but the command you wrote 
# above should return TRUE. 



# How many people in this data set voted for Obama? 
# That is, for how many observations does the variable Voted_For_Obama
# take the value of TRUE? Write R commands that gives you the answer. 
# Note that I know the answer is 4. What you need to give me is the command
# that gives us the answer 4.

# Hint 1: there is a function called length that returns the length of 
#       a vector. One way to do this task is to measure the length of 
#       a subset of the Voted_For_Obama vector where the values are 
#       TRUE. 

# Hint 2: there is a function called nrow that returns the number of 
#       rows of a matrix or a data frame. Another way to do this task 
#       is to measure the number of rows of a subset of the vote.data
#       object for which Voted_For_Obama is equal to TRUE.

```{r}
number_voted_for_obama <- nrow(subset(voter_data, voted_for_obama == TRUE))
number_voted_for_obama
```


# There is one person in this data set who identifies himself as "Ind"
# (Independent). Did he vote for Obama?
#
# Again, I know he didn't. Give me the command that gives us the answer
# FALSE

```{r}
voter_data$voted_for_obama[voter_data$party_id == "ind"]

subset(voter_data, party_id == "ind")$voted_for_obama
```

# Create a subset of the data set that contains only "white" people. 
# Store this smaller data set into an object called vote.white

```{r}
vote.white <- subset(voter_data, race == "White")
```


# Show the third column of the newly created data set vote.white

```{r}
vote.white[3] 
```


# How many white voters in this mini data set voted for Obama? 
# Again, I know the answer is 2. Write a command that gives the answer 2.

```{r}
sum(vote.white$voted_for_obama)
# This counts just the number of TRUEs within a logical variable by treating
# TRUE as 1 and FALSE as 0
```



##### Some extra exercises ######

# Modifying a data frame object -------------------------------------------

# We have created a data frame called my.data in gv900-week5-Review.R.
# Let's see how we can add variables to an existing data frame. 


# To add a new variable, you also use the $ symbol. 
# Specifically, we write DATAFRAME $ NEW_VARIABLE_NAME <- VALUES

# For example, in order to add a new variable called Population 
# that contains the following values: 
# 318946000, 64105654, 127090000, 1367420000, 203322000, 80781000, 87354300

# We write

my.data $ Population <- c(318946000, 64105654, 127090000, 1367420000, 
                          203322000, 80781000, 87354300)

my.data


# If the Console window is wide enough, it should be showing up like this:
#   > my.data
# Country_ID   Country_Name  Regime_Type GDP_PC EU_Member Population
# 1          1  United States    Democracy  51163     FALSE  318946000
# 2          2 United Kingdom    Democracy  39367      TRUE   64105654
# 3          3          Japan    Democracy  46838     FALSE  127090000
# 4          4          China Dictatorship   6070     FALSE 1367420000
# 5          5         Brazil    Democracy  11347     FALSE  203322000
# 6          6        Germany    Democracy  41376      TRUE   80781000
# 7          7          Egypt Dictatorship   3115     FALSE   87354300

# We can see that a new column is now added at the end (far right).


# If the Console window is not wide enough, it may show up like this:
#   
#   > my.data
# Country_ID   Country_Name  Regime_Type GDP_PC EU_Member
# 1          1  United States    Democracy  51163     FALSE
# 2          2 United Kingdom    Democracy  39367      TRUE
# 3          3          Japan    Democracy  46838     FALSE
# 4          4          China Dictatorship   6070     FALSE
# 5          5         Brazil    Democracy  11347     FALSE
# 6          6        Germany    Democracy  41376      TRUE
# 7          7          Egypt Dictatorship   3115     FALSE
# Population
# 1  318946000
# 2   64105654
# 3  127090000
# 4 1367420000
# 5  203322000
# 6   80781000
# 7   87354300

# You may want to adjust the width and ask R to show it again. 

my.data

# If you want to browse a data frame object, use the View function

View(my.data)


# We can create a new variable that is an answer to some operations. 
# For example, GDP_PC measures per capita GDP (in 2013 US dollars). 
# This was calculated as a country's GDP divided by its population:
#   GDP_PC = GDP / Population
# Therefore, if we multiply GDP_PC and Population, we can obtain its
# GDP: 
#     GDP = GDP_PC * Population


# Create a variable within the my.data object called GDP which is equal to
# the product of GDP_PC and Population (GDP_PC times Population).

world$gdppcap08

Write your commands here



my.data


# It is a little bit difficult to read numbers like 1.631823e+13, which 
# means 1.631823 * 10^13. Let's create another variable that shows
# re-scaled GDPs by dividing the raw GDP by 1000000 (on million). 

my.data $ GDP_mil <- my.data $ GDP / 1000000

my.data


# So, now this new variable GDP_mil is shown in 1 million dollars. 
# For example, the value of GDP_mil is 16318234.2 for the United
# States, which means that GDP of the US is 16318234.2 million dollars. 


# Create a new variable within my.data called Is_Democracy that 
# is a logical vector that tells us whether or not a country is democratic. 

# Hint: utilize the Regime_Type variable included in the my.data
#       object. 


Write your commands here



# End of file
